---
header-includes:
   - \usepackage{color}
title: \textcolor{red}{eDentity Sequencer Qc}

date: "`r format(Sys.time(), '%d/%m/%y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,  message=FALSE,warning=FALSE)
```


## 1.1.1 Percantage of bases >= Q30

Metric: Calculate the fraction of bases whose Phred score are greater or equal to Q30.

```{r}

# function to fix misssing commas
fix_csv_commas <- function(input_file, output_file) {
  #  Read the CSV file line by line
  lines <- readLines(input_file)
  
  # Fix lines that have missing commas
  fixed_lines <- sapply(lines, function(line) {
    num_commas <- length(strsplit(line, ",")[[1]]) - 1
    
    # Assuming each row should have 2 commas (for 3 columns)
    if (num_commas == 1) {
      return(paste0(line, ","))   # Add one missing comma
    } else if (num_commas == 0) {
      return(paste0(line, ",,"))  # Add two missing commas
    } else {
      return(line)  # If correct, no change
    }
  })
  
  # Write the fixed lines to the output file
  writeLines(fixed_lines, output_file)
  
  # Return confirmation or any additional actions if needed
  message("CSV has been fixed and saved to: ", output_file)
}

# Example usage:
# fix_csv_commas("input_file.csv", "fixed_file.csv")

#fix missing comma

fix_csv_commas("../results/elements-1.1.1-qc.txt", "../results/elements-1.1.1-qc-fixed.csv")
fix_csv_commas("../results/illumina-1.1.1-qc.txt", "../results/illumina-1.1.1-qc-fixed.csv")


```


```{r}

#extract average q30 fraction
get_mean_q30 <- function(q30_file){
    # read the data
    df <- read.csv(q30_file, header = F)
    
    colnames(df)[1:3] <- c("file", "n_reads", "q30.fraction")
    
    df <- subset(df, select = c("file", "n_reads", "q30.fraction"))
    df <- na.omit(df)
    df$scaled_q30 <- df$n_reads * df$q30.fraction
    
    sum_scaled <- sum(df$scaled_q30) / sum(df$n_reads)
    
    
    return(sum_scaled)
}


q30_df<- data.frame(
        Illumina = get_mean_q30("../results/illumina-1.1.1-qc-fixed.csv"),
        Elements = get_mean_q30("../results/elements-1.1.1-qc-fixed.csv"))

knitr::kable(q30_df,  align="lr", digits = 2)
```



## 1.1.2 Read Length Retension after Q30 Trimming (Percentage).

Metric: (Average length after trimming / original read length ) * 100

```{r}
get_length_retention <- function(untrimmed_stast_file, trimmed_stats_file) {
  
  
  untrimmed_len_stats <- read.csv(untrimmed_stast_file, sep = "\t")
  trimmed_len_stats <- read.csv(trimmed_stats_file, sep = "\t")
  
  return(
    mean(as.numeric(na.omit(trimmed_len_stats$avg_len))) / 
      mean(as.numeric(na.omit(untrimmed_len_stats$avg_len))) * 100
    )
}


length_df <- data.frame(Illumina = get_length_retention(
                                      "../results/illumina-stats.tsv",
                                      "../results/trimmed-illumina-stats.tsv"),
                        
                        Elements = get_length_retention(
                                      "../results/elements-stats.tsv",
                                      "../results/trimmed-elements-stats.tsv"))

knitr::kable(length_df,align="lr", digits = 2)
```


## 1.2.1 Hommopolymer Accuracy

Metric: Average Q score decay per additional base in homopolymer runs (e.g AAAA, TTTT)

```{r}
decay <- function(decay_file){
  
  accuracy <- read.csv(decay_file, header = FALSE)
  colnames(accuracy) <- c("file", "total_reads", "average_decay")
  return(
    mean(accuracy$average_decay)
    )
}

decay_df <- data.frame(
  Illumina = decay("../results/illumina-1.2.1-homopolymer_decay.csv"),
  Elements = decay("../results/elements-1.2.1-homopolymer_decay.csv")
)

knitr::kable(decay_df,align="lr", digits = 2)
```
## 1.2.2 Demultiplexing Efficiency (Percentage)

Metric: (Number of reads successfully demultiplexed / Total number of reads)* 100

N/B: The data from Elements was missing unsuccessfully demultiplexed reads thus we could not calculate 
its demultiplexing efficiency


|       | **Illumina** |  **Element** |
|-------|---------------|----------|
| Total Reads| 885,110,280  |   |
| Demultiplexed Reads| 608,699,786 | 361,464,384 (?) |
| Demultiplexing Efficiency| 68.77 | NA |


```{r}
# Issue:
  # How to get the undetermined in elements: empty with > 1000 reads ??

demux <- read.csv("../results/illumina-1.2.2-demux.csv")

total_reads <- sum(demux$Nreads) # should empty be part of totals ?


n_demultiplexed <-  sum(demux[demux$Category != "Undetermined",]$Nreads)

efficiency <- n_demultiplexed / total_reads * 100

demult_df <- data.frame(illumina = efficiency)

#knitr::kable(demult_df, align="lr", digits = 2)


```

\newpage
## 1.2.3 PhiX Control performance

Metric: Assess sequencing quality using spiked-in Phix control


|       | **Illumina** |  **Element** |
|-------|---------------|----------|
|PhiX Error Rate| 0.005588 | 0.022586|
|PhiX Alignment Rate| 0.2520 |  0.0000 |
|**PhiX Coverage Uniformity:** |    |   |
|     - Coefficient of Variation: | 0.1554 | 0.4593|
|     - Percentage within ±20% of mean coverage: | 0.9332 |0.4330|
| **Additional Information:** |   |   |
|     - Total Reads: | 885,110,280 | 361,464,384 |
|     - Mapped Reads: | 223,075,671 | 482 |
|     - Mean Coverage: | 12,410,202.48 | 26.73 |
|     - Coverage Range: | 12,54,247 - 14,897,715 | 1 - 70 |


## 1.2.4 Duplicate Read Rate (Percentage)

Metrics: Percentage of duplicate reads in a standard non-amplified library.

Only reads mapping to Phix were used here under the assumption that Phix is a "standard non-amplified library"

but it appears Phix reads from Illumina were amplified therefore they may not be a good measure of duplication rate.

|       | **Illumina** |  **Element** |
|-------|---------------|----------|
|Mapped Reads: | 223,075,671 | 482 |
|Duplicate reads:| 219,577,678 | 27 |
| Duplicate rate |  0.9843 |   0.0560 |

## 1.2.5 GC Bias 

Metric: Deviation of coverage in GC-rich regions compared to AT-rich regions

Only reads mapping to Phix were used to compute GC Bias



```{r}
gc_bias <- function(gc_bias_file){
    gc_df <- read.csv("../results/elements-1.2.5.csv")
    return(
     mean(gc_df$Average.Coverage))
  }

gc_bias_coverage = data.frame(
    Elements = gc_bias("../results/elements-1.2.5.csv")
    # illumina coming
)

knitr::kable(gc_bias_coverage,align="lr", digits = 2)
```









